### 2021-08-30 chapters/abstract.tex
add : my old files

\thispagestyle{empty} % this line to make the hide the page number of the first page of the abstract. it is a UCCS requirement

this is the abstract of my thesis