### 2021-09-03 chapters/testing/testing.tex

\newpage

\chapter{Testing}
\label{chapter:testing}

In order to optimize the energy consumption of software we first want to setup a experemental enviroment. This chapter will be dedicated to provide a set of guidelines to provide a \em{reproducible, accurate} and \em{representative} tests.
Each aspect will be detailed in the sections below.

\import{chapters/testing/}{reproducibility}
\import{chapters/testing/}{accuracy}
\import{chapters/testing/}{representativeness}