### 2021-10-21 chapters/testing/testing.tex

\newpage

\chapter{Empirical Evaluation Protocol}
\label{chapter:testing}

In order to optimize the energy consumption of software we first want to setup a experemental enviroment. We have chose to use the empirical appraoch so we will reduce the energy consumption with trial and error.

%  This chapter will be dedicated to provide a set of guidelines to provide a \em{reproducible, accurate} and \em{representative} tests. Each aspect will be detailed in the sections below.

\import{chapters/testing/}{challenges}
\import{chapters/testing}{stateoftheart}
\import{chapters/testing/}{environement}
\import{chapters/testing/}{experement}
\import{chapters/testing/}{conclusion}