### 2021-10-29 chapters/programming_languages/challanges.tex
add : rpc figures

The ultimate goal of this chapter is to provide a guideline on which programming language the developers should chose based on the charecteristics of the project in order to minimize the energy foot print of their product.
No answer is evidedent for a such question. However, we can extract some feartures of each programming langues such as
\begin{itemize}
    \item performance
    \item community support
    \item scalability
    \item energy consumptiom
    \item memory usage
    \item etc
\end{itemize}

first we will start but analysing the behaviour of the general purpose programming langauges with some micro benchkaks, principally for the CLBG game and others from rosetta code base

.