### 2021-11-02 chapters/execution_environement/execution_environement.tex
adding : grpc

\newpage

\chapter{Execution enviroment}
\label{chapter:enviroment}
After studying the impact of the programming languages choice on the energy consumption of softwares. we wanted to dig in more deeper. Therefore we have chosen two of the most popular programming lanaguages \textbf{Python} and \textbf{Java}
In this chapter we will reduce the energy consumption of those following lanaguages by doing some tuning
\import{chapters/execution_environement/java}{java}