### 2022-05-03 chapters/execution_environement/python/python.tex

\subsection{Introduction}
% reasons why we have chosen python

\subsection{python popularity}

%%%% python is a popular programming language  based TIOBE PYPI   add graphs

\subsection{environements}
in this section we will describe the chosen python interpreter

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/python-implementations-tree}
    \caption{Python interpeters}
    \label{fig:interpreters}
\end{figure}

\subsection{benchmarks}

\subsection{first results}

\subsubsection{micro-benchmarks}

\subsubsection{benchmarks}

\subsection{results and foundings}

\subsection{tools and package}

\subsection{Threads to validity}

\subsection{conclusion}