### 2022-05-18 chapters/execution_environement/execution_environement.tex
Revising Java

\newpage

\chapter{Execution environment}
\label{chapter:enviroment}
After studying the impact of the programming languages choice on the energy consumption of software, we wanted to dig in more deeper.
Therefore, we have chosen two of the most popular programming languages \textbf{Python} and \textbf{Java}.
In this chapter, we explore the reduction of energy consumption of those languages by tuning their configuration.

\import{chapters/execution_environement/java}{java}
\import{chapters/execution_environement/python}{python}