### 2022-06-01 chapters/execution_environement/python/python.tex
add : python loops

\newpage
\chapter{python}

\section{Introduction}

When it comes to developing software systems, one can notice that---except Perl---dynamic programming languages have taken over compiled languages along the last decade in terms of developers popularity (cf. Figure~\ref{fig:pypl}).
However, it remains unclear if this class of dynamic programming languages can reasonably compete with compiled ones when it comes to power efficiency.

\begin{figure}[htbp]
    \includegraphics[width=\linewidth]{imgs/programminglanguangespopularity.png}
    \caption{PYPL Popularity of Programming Languages~\cite{noauthor_pypl_nodate}}
    \label{fig:pypl}
\end{figure}

In particular, Noureddine~\emph{et~al.}~\cite{noureddine_preliminary_2012} in 2012, and then Pereira~\emph{et~al.}~\cite{pereira_energy_2017} in 2017, conducted empirical power measurements on this topic, and both concluded that compiled programming languages overcome dynamic ones when it comes to power consumption.
According to their experiments, an interpreted programming language like Python can impose up to a $7,588\,\%$ energy overhead compared to C~\cite{pereira_energy_2017} (cf. Figure~\ref{fig:hannoi}).
% However, the operational conditions under which these measurements may threaten the validity of their results.

\begin{figure}[htbp]
    \includegraphics[width=\linewidth]{imgs/hannoiimplementation.png}
    \caption{Energy consumption of a recursive implementation of Tower of Hanoi program in different languages~\cite{noureddine_preliminary_2012}}
    \label{fig:hannoi}
\end{figure}
\subsection{python popularity}

However, Python seems to attract a large community of developers interested in data analysis, web development, system administration, machine learning---according to a survey conducted in 2018 by JetBrains---\footnote{\url{https://www.jetbrains.com/research/python-developers-survey-2018/}}one can fear that the wide adoption of dynamic programming languages, like Python, in production may critically hamper the power consumption of ICT.

As the popularity of such dynamic programming languages partly builds on the wealth and the diversity of their ecosystem (\emph{e.g.}, the NumPY, SciKit\,Learn, and Panda libraries in Python), one cannot reasonably expect that developers will likely move to an alternative programming language mostly for energy considerations.
Rather, we believe that a better option consists in leveraging the strength of this rich ecosystem to promote energy-efficient solutions in order to improve the power consumption of legacy software systems.

In this chapter, we therefore explore the oblivious optimizations that can be applied to Python legacy applications in order to reduce their energy footprint.
As Python is widely adopted by software services deployed in public and private Cloud infrastructures, we believe that our contributions will benefit to wide diversity of legacy systems and not only favorably contribute to reduce the carbon emissions of ICT, but also reduce their cloud invoice for the resources consumed by these services.
More specifically, this chapter focuses on runtime optimizations that can adopted by developers to leverage the power consumption of Python applications.
We start by studuing the impact of some of programmers choice such as the type of data structures or loops, on the global energy consumption of the execution code.
then we discuss some other factors such as level of concurrency etc. later we we will talk about another non intrusive ways to optimize the energy consumption of this code. one kind of those optimizations includes alternative interpertersand some libraries that are dedicated to optimize the code without changing its structures such as \emph{ahead-of-time} (AOT) compilation and \emph{just-in-time} (JIT) libraries that are maintained by the community.

% The contributions of this chapter can therefore be summarized as:
% \begin{compactenum}
%     \item a classification of state-of-the-art Python optimizations,
%     \item an evaluation of the power efficiency of these optimizations on well-known micro-benchmarks,
%     \item a comparison of the effective energy efficiency of Python compared to compiled languages,
%     \item an assessment of the energy speedup of these optimizations on representative applications and workloads.
% \end{compactenum}

%%%% python is a popular programming language  based TIOBE PYPI   add graphs

\subsection{python gluttony}

the simplicity, flexibility and the clarity of Python code led this programming language to a huge success. Therefore, a larger community and a wide range of support and libraries that allowed it to touch several fields like data analyses, machine learning, natural language processing, web development, astrophysics etc.
% python is not efficient 
% 1 Energy Efficiency Across Programming Languages: How Do Energy, Time, and Memory Relate?
%  \begin{figure}[hbt]
%     \includegraphics[width=\linewidth]{figures/hannoiimplementation.png}
%     \caption{Energy consumption of the recursive implementation of Tower of Hanoi program in different languages (using a base 10 logarithmic scale) \cite{noureddine_preliminary_2012}}      
%     \label{fig:hannoi}
% \end{figure}

However, and according to \cite{pinto_energy_2017,noureddine_preliminary_2012} and [1]. Python tends to be one more energy hungry . In figure \ref{fig:hannoi}, python consumes 30 times more than C or C++. The test was done on an implementation of \fnurl{hannoi test}{ https://en.wikipedia.org/wiki/Tower_of_Hanoi}  30 disks.

Python consumes a lot of energy mainly because it is slow in execution. Its Flexibility and simplicity caused it to drop off in performances, because Python gains its flexibility from being a dynamic language, therefore it needs an interpreter to execute its programs, which makes them much slower compared to the other ones that are written in compiled programming languages such as C and C++ or semi-compiled languages like Java.

As shown in Table \ref{fig:clbg}, one can see that in the most of the test cases Python takes more time to execute -the only case that he wasn't the last one was in the benchmark regx-redux where he beaten Go-, and in some cases the gab was huge such as in the n-body test where python took around 100 times more than C++.

\begin{table}[hbt]
    \begin{tabular}{l|*{5}c}
                           & C    & C++   & Java & python & Go    \\
        \hline
        pidigits           & 1.75 & 1.89  & 3.13 & 3.51   & 2.04  \\
        reverse-complement & 1.75 & 2.95  & 3.31 & 16.76  & 4.00  \\
        regx-redux         & 1.45 & 1.66  & 10.5 & 15.56  & 28.69 \\
        k-nucleotide       & 5.07 & 3.66  & 8.66 & 79.79  & 15.36 \\
        binary-trees       & 2.55 & 2.63  & 8.28 & 92.72  & 28.90 \\
        fasta              & 1.32 & 1.33  & 2.32 & 62.88  & 2.07  \\
        Fannkuch-redux     & 8.72 & 10.62 & 17.9 & 547.23 & 17.82 \\
        n-body             & 9.17 & 8.24  & 22.0 & 882.00 & 21.00 \\
        spectral-norm      & 1.99 & 1.98  & 4.27 & 193.86 & 3.95  \\
        Mandelbort         & 1.64 & 1.51  & 6.96 & 279.68 & 5.47
    \end{tabular}
    \caption{Comparison of the execution time between different programming languages using the computer language benchmarks game \cite{noauthor_which_nodate}}
    %  TODO : ADD reference 
    \label{fig:clbg}
\end{table}

In this chapter we will target the following research question :
\begin{enumerate}
    \item Do the data structres and controle structures impact the energy consumption ?
    \item how does the parallelism affect the energy behaviour of python programs ?
    \item are they any non intrusive methodes to reduce the imapct of the energy consumption of existing python softwares ?
    \item how can one apply those findings in the real world ?
\end{enumerate}

\section{experimetal protocole}

as we have discussed in the previous chapter \ref{chapter:benchmarking}. instead of running the tests the idea was to design a system that allows practiouners to reproduce and extends our tests. and then we use the same system to run answer some of researchs questions
%%%% probably ill resue the same things from other chapters

\section{python insights}

The goal of this section is to find some insights where we can optimize the energy consumption without impacting the source code. so we started by extending the work of \citeauthor{hasan_energy_2016} and \citeauthor{oliveira_recommending_nodate} to the python environement.

%%%% concurency

Figure \ref{fig:python_multiprocessing} shows the behaviour of python programs when we try to introduce the parallelism.
As we know python is a single threaded program thx to ghe GLS ( global lock system ) howerver due to the increase of the number of cores /threads per cpu it many libraries stared to take advanteage of this feature so most of them they will try to simulate the multithreading by using multiple instances pf the processor % TO VERIFY

\begin{figure}[hbt]
    \centering
    \includegraphics[width=\linewidth]{imgs/multiprocessing_energyvstime}
    \caption{energy behaviour based on multiprocessing}
    \label{fig:python_multiprocessing}
\end{figure}

as we can see in the graph the energy consumption is correlate with the number of threads until we arrive to the limit of the cores and then we lose the adventage of the multiprocessing , well in that case we pass from parallelism to concurency. where different sub process have to compete for the CPU Ressources.
Another finding is when we hit the number of physical cores. there was an increase of execution time but still reduced energy, the reason behind this is the scheduler of the operating system. basicaly he favorits the hyper threads than the physical core wich will lead to some context switches which cause the slow behaviour, howerver in the other hand the other two physical cores are not consuming energy, neither their hyper threads which explains the gain of the eneryg consumption in this case.
in the Chapter %% reference the work of intern from lyon  
we discuss a deeper this behaviour of the scheduler and in that case we confirm that it is not related to python but it is more generic behaviour %% maybe even this graph should go there

another field of investigation is the type data and structure conroles. that might impact the energy consumption.
To do so. We iterate over a list using three methodes.
first the classical for ( for i in range (len(n))). howerver as we can see here unlike other porgramming langauges it requires extra operations such as determining the length of the collection , and then using the iterator range. so we tried the more adapted version
( for element in collection). Morever in most programming languages the for loop is translated to a while loop ( transfirmation from D type to B type -- asm -- ), therefore we wanted to compare this with a ( while ) version.
After determining the main ways to iterate over a loop we run the algorithms  collections of different data types to see wether it will impact of the energy consumption of the code, same thing for the size of the collection.

As we can see in figure~\ref{fig:pythonloops} the type of the data hasn't an impact on the energy consumption.In the other hand the way we iterate over the collection has a huge factor. interestely, the for in range loop was by far the optimal one with following by the regural for in collection, and the while part was the least one with an overhead of 400\% compared to the first option.

%%% putting the idea, later i ll reformulate it 
the reason behind a such behaviour is mainly related on how python interpeter is made.
% add refenece https://www.python.org/doc/essays/list2str/ 
%% https://www.pythonpool.com/for-vs-while-loop-python/  
% https://wiki.python.org/moin/TimeComplexity
%% http://python-history.blogspot.com/2010/06/from-list-comprehensions-to-generator.html?m=1
In order to reduce the latency of the python code. most of the buitl-in functions an operations are written in C, same thing goes for the function range, further more the function len has a complexity of O(1) because it is based on the function Py\_SIZE  of C which stores the length in a field for the object.

so basically for the for in range is basically creating a new iterator that has the same length of the first one. and for each iteration we will have to do a second acceess ( l[i]) instead of one one access hence the double time.

for the wile is even slower due to the implicite incrementation of the variable which will cause an extra operation during the loop .

to confirm the hyphothsis. we tried to construct a new list by editing the elements of the previous one. And as predicted , the builtin methodes are the best energy saving, while the customize while loop is the heaviest. an interesting finding is. the impact of annonymous functions ( lambda expressions) on the energy consumption.
the reseason behind this is the fact that python treats those functions as local variables unlike the predifined ones which are global in our case. therefore they are faster and consumes less energy  % same reference https://www.python.org/doc/essays/list2str/

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_iterations}
    \caption{energy behaviour of different python loops }
    \label{fig:pythonloops}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\linewidth]{imgs/python_treatemens}
    \caption{energy behaviour of different methodes to change a list}
    \label{fig:pythontreatement}
\end{figure}

%% should i include the part with the exceptions ???

\newpage
\subsection{Python Runtimes}

we used the following implementations.

We should explain the methodology we used to select these

\begin{enumerate}
    \item \textbf{CPython}:\footnote{\fnurl{https://www.python.org/}} a Python interpreter written in C programming language. It is the reference interpreter of Python. CPython compiles the source code into a bytecode and then interprets it. The CPython project still supports both versions of Python 2 and 3;
    \item \textbf{PyPy}:\footnote{\url{http://pypy.org}} An alternative implementation of the Python interpreter. It is written using \emph{RPython} in order to use the JIT. It compiles the most used portions of the Python code into a binary code for better performance. To benefit from these optimizations, the program has to be executed for at least for few seconds so the JIT has enough time to warm up, the JIT optimization will be applied only the code written by the programmer and not external libraries;
    \item \textbf{Cython}:\footnote{\url{https://github.com/cython/cython}} a static compiler for Python. It translates the Python code into a C code, and then compiles it using a C compiler. It also support(s) an extended version of the Python language that allows programmers to call \emph{C functions}, declare \emph{C types} and use static types which will help the translation of Python objects into native types, such as integers, float, which often means a better performances, since native C libraries are almost all the time faster than the python written once~\cite{pereira_energy_2017}.
    \item \textbf{Intel\,Python}:\footnote{\url{https://software.intel.com/en-us/distribution-for-python}} A customized interpreter developed by Intel in order to enhance performances of Python programs. It is dedicated to data sciences, high-Performance computing. It uses some Intel kernel libraries, such as Math Kernel Library (Intel MKL\footnote{\url{https://software.intel.com/en-us/mkl}}) and data analytics acceleration library (Intel DAAL\footnote{\url{https://software.intel.com/en-us/intel-daal}}). It supports both versions of Python;
    \item \textbf{Active\,Python}:\footnote{\url{https://www.activestate.com/products/activepython/}} developed by the Activestates company, it provides a standardized Python distribution to ensure license compliance, security, compatibility and performance. Therefore, ActivePython has its own built-in packages (more than 300 packages) and supports both versions of Python;
    \item \textbf{IronPython}:\footnote{\url{https://ironpython.net}} A .Net-based Python interpretation platform written in C\# in order to be used over the .Net vm or mono, it benefits from all the optimizations of .Net virtual machine, such as the JIT and garbage collector mechanisms.
    \item \textbf{GraalPython}:\footnote{\url{https://github.com/graalvm/graalpython/}} A Python interpreter that is based on GraalVM \footnote{\url{https://www.graalvm.org/docs/why-graal/}} (a universal virtual machine developed by oracle for running applications written in different programming languages). For the time being, it only supports Python~3 and it is still in the experimental stage;
    \item \textbf{Jython}:\footnote{\url{https://jython.github.io}} Implementation of Python programming language written in Java for the \emph{Java virtual machine} (JVM). Similar to IronPython and GraalPython, it leverages the optimization mechanisms provided by the JVM to enhance the Python programs;
    \item \textbf{MicroPython}:\footnote{\url{http://micropython.org}} a lightweight Python version dedicated to embedded systems and micro-controllers;
    \item \textbf{Nuitka}:\footnote{\url{http://nuitka.net/pages/overview.html}} a Python compiler written in Python that generates a binary executable from Python code. It translates the Python code into a C program that is then compiled into a binary executable;
    \item \textbf{Numba}:\footnote{\url{https://numba.pydata.org}} a library that includes JIT compiler in order to enhance the performances of Python functions using the industry-standard LLVM compiler library;
    \item \textbf{Shedskin}:\footnote{\url{https://github.com/shedskin/shedskin}} a static transpiler that translates implicitly statically typed python into C++ code.
    \item \textbf{Hope}~\cite{akeret_hope_2015}: a Python library that aims to introduce JIT compiler into the Python code;
    \item \textbf{Parakeet}~\cite{DBLP:conf/hotpar/RubinsteynHWS12}: a runtime accelerator for an array-oriented subset of Python;
    \item \textbf{Stackless\,Python}:\footnote{\url{https://github.com/stackless-dev/stackless/wiki}} an interpreter that focus on enhancing multi-threading programming.
    \item \textbf{Pyjion}:\footnote{\url{https://github.com/microsoft/pyjion}} JIT API for CPython, same purpose as Parakeet and Hope.
    \item \textbf{Pyston}:\footnote{\url{https://blog.pyston.org}} performance-oriented Python implementation built using LLVM and modern JIT techniques. the project is funded by dropbox;
    \item \textbf{Grumpy}:\footnote{\url{https://github.com/google/grumpy}} a source-to-source transpiler that translates the Python code into a Go code that will be compiled to a binary executable. It also offers an interpreter called \emph{grumprun} which can directly execute the Python code. Unfortunately, we cannot use it because the project is already outdated (last commit is in 2017) and it has a lot of limitation in term of supporting the Python language, such as some built-in functions and standard libraries;
    \item \textbf{Psyco}:\footnote{\url{http://psyco.sourceforge.net}} a JIT compiler for Python;
    \item \textbf{Unladen\,Swallow}:\footnote{\url{https://unladen-swallow.readthedocs.io/en/latest/}} an attempt to (use) LLVM  as  JIT compiler for CPython.
\end{enumerate}

\subsection{Runtime Classification}
Before futher proceeding with the list of candidate runtimes for Python applications, we propose a classification according to several criterions:
\begin{description}
    \item[\bf Type] refers to the category of runtime infrastructure that supports the execution of a Python application. In particular, we consider 3 types of environments: \emph{Interpreter}, \emph{Compiler} and \emph{Library}. \emph{Interpreter} refers to the class of environment that does not require any preprocessing of Python source code. \emph{Compiler} introduces a compilation phase prior to the execution of the application. Finally, \emph{Library} induces some modification of the source code.
    \item[\bf Runtime] refers to the technology supporting the execution of a Python application. This technology can refer to the programming language used to program the interpreter, the target language for a compiler or a library.
    \item[\bf JIT optimisation] refers to the support of \emph{just-in-time} compilation in the runtime infrastructure supporting the execution of the application.
    \item[\bf GC optimisation] refers to the support of \emph{garbage collection} in the runtime infrastructure supporting the execution of the application.
    \item[\bf Python version(s)] refers to the list of Python source code versions supported by the runtime environment.
\end{description}

We should explain the classification of these runtimes

\begin{table}
    \caption{Classification of Python implementations}
    \label{fig:python-classes}
    \center
    \begin{tabular}{|l|c|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{\bf Name} & \multirow{2}{*}{\bf Type} & \multirow{2}{*}{\bf Runtime} & \multicolumn{2}{c|}{\bf Optimisations} & \multicolumn{2}{c|}{\bf Python}                     \\
        % \cline{1-5} 
                                  &                           &                              & {\bf JIT}                              & {\bf GC}                        & {\bf 2} & {\bf 3} \\
        \hline
        \hline
        CPython                   & Interpreter               & C                            & \no                                    & \no                             & \yes    & \yes    \\
        \hline
        Intel\,Python             & Interpreter               & C                            & \no                                    & \no                             & \yes    & \yes    \\
        \hline
        ActivePython              & Interpreter               & C                            & \no                                    & \yes                            & \yes    & \yes    \\
        \hline
        PyPy                      & Interpreter               & Python                       & \yes                                   & \yes                            & \yes    & \yes    \\
        \hline
        IronPython                & Interpreter               & .Net                         & \yes                                   & \yes                            & \yes    & \yes    \\
        \hline
        GraalPython               & Interpreter               & GraalVM                      & \yes                                   & \yes                            & \no     & \yes    \\
        \hline
        Jython                    & Interpreter               & Java                         & \yes                                   & \yes                            & \yes    & \no     \\
        \hline
        Stackless\,Python         & Interpreter               & python                       & \no                                    & \no                             & \yes    & \no     \\
        \hline
        MicroPython               & interpreter               & c                            & \no                                    & \no                             & \no     & \yes    \\
        \hline
        Pyston                    & interpreter               & LLVM                         & \yes                                   & \no                             & \yes    & \no     \\
        \hline
        Unladen\,Swallow          & Interpreter               & LLVM                         & \yes                                   & \no                             & \yes    & \no     \\
        \hline
        Cython                    & Compiler                  & C                            & \no                                    & \no                             & \yes    & \yes    \\
        \hline
        Nuitka                    & Compiler                  & C                            & \no                                    & \no                             & \yes    & \yes    \\
        \hline
        Shedskin                  & Compiler                  & C++                          & \no                                    & \no                             & \yes    & \yes    \\
        \hline
        Grumpy                    & Compiler                  & Go                           & \no                                    & \no                             & \yes    & \yes    \\
        \hline
        Numba                     & Library                   & C                            & \yes                                   & \no                             & \yes    & \yes    \\
        \hline
        Hope                      & Library                   & python                       & \yes                                   & \no                             & \yes    & \yes    \\
        \hline
        Psyco                     & Library                   & python                       & \yes                                   & \no                             & \yes    & \yes    \\
        \hline
        Pyjion                    & Library                   & .NET Core                    & \yes                                   & \no                             & \yes    & \yes    \\
        \hline
        Parakeet                  & library                   & C                            & -                                      & \no                             & \yes    & \no     \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[hbt]
    \caption{Classification of Python implementations}
    \label{fig:pythonimplementations}
    \center
    \begin{tabular}{|l|c|c|c|}
        \hline
        Version  & Interpreter     & Transpiler/Compiler & Jit library \\
        \hline
        \hline
        Python~2 & Cpython2        & Cython2             & Numba 2     \\
        % \cline{2-4} 
                 & Pypy2           & Shesdskin           & Hope        \\
        % \cline{2-4} 
                 & Pytson          & Grumpy              & Parakeet    \\
                 & Ironpython      &                     & Psyco       \\
                 & Jython          &                     & Pyjion      \\
        % &  &  &  \\
                 & Micropython     &                     &             \\
                 & Pysec           &                     &             \\
                 & StacklessPython &                     &             \\
        % \cline{2-4} 
        \hline
        Python~3 & Cpython3        & Nuitka              & Numba3      \\
                 & Pypy3           &                     &             \\
                 & GraalPython     &                     &             \\
        \hline
    \end{tabular}
\end{table}

There are other implementations that we didn't considerate because either the project aborted many years ago or it has a very limited support for python features.
After the collection of those implementation we filtered them. in order to keep only the versions that are still maintained and support most of python features.
and we classified them into 3 categories depending of their integration with the python code.  In the \ref{fig:pythonimplementations} we describe the implementations that we kept and the version of each implementation and its category.

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/python-implementations-tree}
    \caption{Python interpeters}
    \label{fig:interpreters}
\end{figure}

\subsection{tests preparation}

To study the behaviour of the python implementation regarding the energy consumption, we have to focus on the effect of the implementation and mitigate the maximum any side effects such as the organisation of the code or any extra consumption due to the operating system or tier libraries.
Therefore, for each test we took the implementation written in python2 as a reference and tried to use it in other implementations as it is. If it is not supported by python3, we transformed the code using the officiel library  \emph{2to3}\footnote{\url{https://docs.python.org/3.7/library/2to3.html}}.
In the case of the libraries that use \emph{JIT} adding a decorator to the function that we want to optimize was enough, if there are other changes we assume that they alter the original code which is against our purpose.

Each test is implemented in a Docker container for the several reasons
\begin{itemize}
    \item Isolation: each container has only the test program implemented with a single python runtime to remove any interference between different implementations.
    \item Deplpoyment: to use the testing machine without extra configurations that may alter the behaviour of the os toward the enrgy consumtion
    \item Reproductibility: One of the most frequent benchmark crimes~\cite{DBLP:journals/corr/abs-1801-02381} in research is the lack of Reproductibility, by using Docker we ensure that each test has an Image that will be accessible in public.
\end{itemize}

Despite the presence of the official docker images for the most of the runtimes, we prefered to build our own using the same reference image in order to remove any bias due to the Os used in the offical image.We used ArchLinux with the kernel version 4.9.184 as a base image.

\begin{table}[hbt]
    \begin{tabular}{ll}
        \hline
        CPU     & Intel Xeon Gold 6130 (Skylake, 2.10GHz, 2 CPUs/node, 16 cores/CPU)                                                     \\
        Memory  & 192 GiB                                                                                                                \\
        Storage & 240 GB SSD SATA Samsung MZ7KM240HMHQ0D3                                                                                \\
                & 480 GB SSD SATA Samsung MZ7KM480HMHQ0D3                                                                                \\
                & 4.0 TB HDD SATA Seagate                                                                                                \\
        Network & eth0/enp24s0f0, Ethernet, configured rate: 10 Gbps, model: Intel Ethernet Controller X710 for 10GbE SFP+, driver: i40e \\
                & ib0, Omni-Path, configured rate: 100 Gbps, model: Intel Omni-Path HFI Silicon 100 Series [discrete], driver: hfi1      \\
        \hline
    \end{tabular}
    \caption{Testing Machine Configuration }
    \label{fig:dahuconfig}
\end{table}

\subsection{experimental setup}

all our tests have been executed in a Dell PowerEdge C6420 server machine. A summary of its hardware is listed in table \ref{fig:dahuconfig}. The machine is equiped with a minimal version of Debian\,9 (4.9.0 kernel version) where we install Docker (version 18.09.5).

\begin{figure}[hbt]
    \includegraphics[width=\linewidth]{imgs/SmartWatts.png}
    \caption{powerapi architecture}
    \label{fig:powerapi}
\end{figure}

\subsection{Power Measurements}

As we know, the energy of a program is the integrale of its power overtime. For us ower case, we used Intel \emph{Running Power Average Limit} (RAPL)~\cite{Khan:2018:RAE:3199681.3177754} to collect the power samples of the running tests,

We used used \textsc{PowerAPI}~\cite{DBLP:journals/jss/ColmantRKSFS18}, to report Data collected by intel Rapl and send it into another machine that we call computing machine. then we calculate the Energy using  the the trapezoidal rule.

\ref{fig:trapezrule}
\begin{figure}[hbt]
    \centering
    \begin{equation}
        E = \int^a_b P(t)dt \simeq \sum^n_{k=1} \frac{P(t_k-1)+P(t_k)}{2}
    \end{equation}
    \caption{}
    \label{fig:trapezrule}
\end{figure}

figure \ref{fig:powerapi} shows the architecture of our testing model.

The reason of separation between data collection and energy calculation is to minimise any interference with the test so our sensor is a light c program running inside a docker container.

\begin{figure}
    \caption{energy consumption of different impelementations using Bit Operation benchmarks (Joule) }
    \label{table:bitops}
    \begin{tabular}{lrrrrr}
        \toprule
        benchmark    & A          & GG         & LL         & Or         & XOR        \\
        \midrule
        activepython & 676.980672 & 763.208040 & 651.783048 & 743.016425 & 728.828481 \\
        cpython2     & 441.082298 & 435.886950 & 430.846171 & 415.247383 & 419.081447 \\
        cpython3     & 595.209019 & 685.085315 & 563.839300 & 657.972734 & 655.560574 \\
        cython       & 35.077408  & 182.688395 & 274.177830 & 34.868014  & 34.504778  \\
        nuitka       & 33.260991  & 32.980380  & 33.256450  & 33.472770  & 33.030889  \\
        numba2       & 9.102939   & 8.411176   & 9.460620   & 9.375920   & 9.755952   \\
        numba3       & 9.566646   & 10.144212  & 9.219703   & 9.344613   & 9.665108   \\
        pypy2        & 8.456867   & 7.844918   & 8.286278   & 8.138692   & 7.952999   \\
        pypy3        & 7.552731   & 8.093884   & 8.108508   & 8.669448   & 8.623737   \\
        shedskin     & 8.024198   & 8.070870   & 8.399917   & 8.126236   & 8.277546   \\
        \bottomrule
    \end{tabular}

\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/barplot_binarry_tree}
    \caption{energy behaviour based on multiprocessing}
    \label{fig:python_multiprocessing}
\end{figure}

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/bitopts_mean}
    \caption{Mean consumption of different implementations of bit operations (Joule) }
    \label{fig:bitops}
\end{figure}

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/alltomti_performance}
    \caption{different interpererts optimisation }
    \label{fig:tommi_all}
\end{figure}
\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__pypy2_pypy3}
    \caption{green factor of pypy }
    \label{fig:pypy2vspypy3}
\end{figure}

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__cpython3_numba3_pypy3}
    \caption{comparaison of pypy vs python vs numba }
    \label{fig:p3}
\end{figure}

\begin{figure}[thb]
    \centering
    \includegraphics[width=\linewidth]{imgs/tommti_compare__cpython2_pypy2_numba2}
    \caption{comparaison of pypy vs python vs numba }
    \label{fig:p2}
\end{figure}

\subsubsection{micro-benchmarks}

\subsubsection{benchmarks}

\subsection{results and foundings}
As we can see in figure \ref{fig:tommi_all}, there is no evolution between cpython2 and cpython3
Intelpython and active python both folows the same behaviour. one can conclude that the work that have been on those interpreters is mainly
to improve a specific purpose, Active python claims that their version is focused on secuirty which explains the lack of some performances due to the introduction of more reonforcment
.Morver Intel published their version of python as a dedicated for machine learning. Unfortunately the tommti benchmark is a set that focuses on the general purpose programming
which does not reflect the performance of the machine learning.
Another aspect of a such behaviour might be due to the fact of the processors that were used in the testbed. and it might change in the future if we include the GPU part for the machine learning benchmarks
% reference the work of the intern 
As for nuitka. there were no optimization in the energy consumpotion despite the fact that is it a complier.
However if we dig through the nuitka mechanisms. they basically embed the python code with an interpreter.
Unllike nuitka shedskin exhebit a the best energy consumption pattern when it comes to the arithmetic operartions. One can conclude it is due to the fact of the native type of the variables. unlike the interpreters where they are treated as object in the begining.

for the other interpeters pypy is very promising especially when it comes to data maniupulation as one can see in the figure \ref{fig:bar_tommti_vector}
pypy is by far the best interperer when it comes to treating vectors.
numba2 introduced the JIT but wasn't as promising as numba3.

for the other vm based interpereters. jython and ipy lacked in term of energy optimisation which was kinda expected since they were in their the begining of the stage and the main puprose of such implementation is to link the bytecode generated by jython and ironpython with their respectives virtual machin.

Unlike the the previous interpeters. graal exhebits a certin promises when it comes to complex algorithms - nested loops -
micro python is dedicated to embeded systems so lunching it is powerful cpu machines will be misleading.

Most of the interperters had the same behaviour when it comes to the input outputs. except for jython which was kinda of a anomaly probably due to the lack of the optimizations.
% pypy handles exceptions very well

\subsection{use cases}
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/django}
    \caption{energy behaviour based on multiprocessing}
    \label{fig:django}
\end{figure}

%% todo 
% add performance 
% add dram \ a glimps 
% add bitops

\subsection{tools and package}

\subsection{Threads to validity}

\subsection{conclusion}