### 2022-07-02 chapters/benchmarking/benchmarking.tex
refactoring

\newpage
\label{chapter:benchmarking}
\chapter{Benchmarking Protocol for measuring Energy Consumption of stoftwares}

\import{\currfiledir}{draft}

\import{\currfiledir}{reproducibility}

\import{\currfiledir}{experiment}
\import{\currfiledir}{conclusion}

% However for them to be more representatives we wanted to simulate the production enviromenet.
%% should i add  the thing about green faas ? 
%% same about what we have done with python workers and lazy ? basically just a graph