### 2022-07-02 chapters/benchmarking/reproducibility.tex
fix : bug where i cant add references

%  State of the art.

% - add : Evaluate Collaboratory article 
% - add : the  hardware papers

%% this is might be the part that treats reproducibility for my case