### 2022-09-03 chapters/abstract.tex

% ************************** Thesis Abstract *****************************
% Use `abstract' as an option in the document class to print only the titlepage and the abstract.
\begin{abstract}
    This is where you write your abstract \ldots
\end{abstract}