### 2022-09-22 chapters/benchmarking/benchmarking.tex
update :benchmarking protocol

\chapter{A Benchmarking Protocol to Measure Software Energy Consumption}\label{chapter:benchmarking}
\section{Introduction \note{missing} }
This chapter covers ways to overcome empirical analysis challenges in energy consumption studies. First, we go through the three components of a successful benchmark within the energy consumption experiment. Section~\ref{sec:benchmarking_reproducibility} concentrates on the "reproducibility" issue and how to deliver reproducible experiments without interfering with the energy measurement,while Section~\ref{sec:taming-the-energy-variation} discusses accuracy in software energy consumption.
\import{\currfiledir}{draft}
\import{\currfiledir}{experiment}
\import{\currfiledir}{conclusion}

% However for them to be more representatives we wanted to simulate the production enviromenet.
%% should i add  the thing about green faas ? 
%% same about what we have done with python workers and lazy ? basically just a graph