### 2022-09-23 chapters/benchmarking/benchmarking.tex
Proofreading titles

\chapter{Benchmarking Protocol to Measure Software Energy Consumption}\label{chapter:benchmarking}

\import{\currfiledir}{draft}
\import{\currfiledir}{experiment}
\import{\currfiledir}{conclusion}

% However for them to be more representatives we wanted to simulate the production enviromenet.
%% should i add  the thing about green faas ? 
%% same about what we have done with python workers and lazy ? basically just a graph