### 2022-10-18 chapters/benchmarking/benchmarking.tex

\chapter{Benchmarking Protocol to Measure Software Energy Consumption}\label{chapter:benchmarking}
% \section{Introduction \note{missing} }
This chapter covers ways to overcome empirical analysis challenges in energy consumption studies.
First, we go through the three components of a successful benchmark when performing energy-related experiments.
Section~\ref{sec:benchmarking_reproducibility} focuses on the "reproducibility" challenge---to deliver reproducible experiments without interfering with the energy measurement---while Section~\ref{sec:taming-the-energy-variation} discusses the \emph{accuracy} of software energy consumption.

\import{\currfiledir}{draft}
\import{\currfiledir}{experiment}
\import{\currfiledir}{conclusion}

% However for them to be more representatives we wanted to simulate the production enviromenet.
%% should i add  the thing about green faas ? 
%% same about what we have done with python workers and lazy ? basically just a graph