### 2022-10-18 chapters/programming_languages/rpc/tail_latency_clients.tex

\begin{tabular}{lllrrrrrrrrrrrr}
{} & {} & {payload} & {1005} & {905} & {805} & {705} & {605} & {505} & {405} & {305} & {205} & {105} & {5} & {1} \\
{languageGroup} & {languageType} & {name} & {} & {} & {} & {} & {} & {} & {} & {} & {} & {} & {} & {} \\
\multirow[c]{5}{*}{1} & \multirow[c]{5}{*}{compiled} & cpp grpc st & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 15.0%, transparent 15.0%) 1.752 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 13.4%, transparent 13.4%) 1.724 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 11.2%, transparent 11.2%) 1.900 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 10.8%, transparent 10.8%) 1.842 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 11.3%, transparent 11.3%) 1.898 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 10.8%, transparent 10.8%) 1.843 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 12.6%, transparent 12.6%) 2.038 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 10.6%, transparent 10.6%) 1.811 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 11.0%, transparent 11.0%) 1.844 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 9.0%, transparent 9.0%) 1.881 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 1.804 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 5.359 \\
 &  & crystal grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 27.5%, transparent 27.5%) 2.998 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 28.4%, transparent 28.4%) 3.261 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 23.9%, transparent 23.9%) 3.233 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.2%, transparent 26.2%) 3.415 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 24.4%, transparent 24.4%) 3.244 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.7%, transparent 26.7%) 3.457 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 31.8%, transparent 31.8%) 4.009 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 42.8%, transparent 42.8%) 5.055 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 67.0%, transparent 67.0%) 7.433 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 13.167 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 202.268 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 754.372 \\
 &  & go grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 5.0%, transparent 5.0%) 0.758 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.1%, transparent 4.1%) 0.766 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.761 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.763 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.768 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.768 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.770 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.772 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.775 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.808 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 2.056 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 7.307 \\
 &  & rust tonic st & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 20.9%, transparent 20.9%) 2.337 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 19.2%, transparent 19.2%) 2.313 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 16.0%, transparent 16.0%) 2.407 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 14.4%, transparent 14.4%) 2.208 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 13.1%, transparent 13.1%) 2.087 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 12.9%, transparent 12.9%) 2.051 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 11.6%, transparent 11.6%) 1.939 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 12.5%, transparent 12.5%) 2.003 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 12.4%, transparent 12.4%) 1.985 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 9.1%, transparent 9.1%) 1.896 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 2.037 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 5.788 \\
 &  & swift grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 38.4%, transparent 38.4%) 4.086 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 41.3%, transparent 41.3%) 4.590 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 33.3%, transparent 33.3%) 4.208 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 44.2%, transparent 44.2%) 5.257 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 47.3%, transparent 47.3%) 5.605 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 46.1%, transparent 46.1%) 5.421 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 45.7%, transparent 45.7%) 5.439 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 48.4%, transparent 48.4%) 5.617 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 48.7%, transparent 48.7%) 5.610 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 38.3%, transparent 38.3%) 5.515 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.9%, transparent 1.9%) 5.445 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.4%, transparent 0.4%) 8.337 \\
\multirow[c]{2}{*}{2} & \multirow[c]{2}{*}{vm .net} & csharp grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 7.2%, transparent 7.2%) 0.981 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 6.2%, transparent 6.2%) 0.978 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.1%, transparent 2.1%) 0.953 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.1%, transparent 2.1%) 0.948 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.9%, transparent 1.9%) 0.937 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.8%, transparent 1.8%) 0.927 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.8%, transparent 1.8%) 0.935 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.0%, transparent 2.0%) 0.938 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.2%, transparent 2.2%) 0.959 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.0%, transparent 2.0%) 1.019 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.6%, transparent 0.6%) 2.962 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.7%, transparent 0.7%) 10.767 \\
 &  & dotnet grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 5.2%, transparent 5.2%) 0.775 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.3%, transparent 4.3%) 0.786 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.773 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.768 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.766 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.765 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 0.759 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.759 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 0.756 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 0.774 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.4%, transparent 0.4%) 2.594 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.4%, transparent 0.4%) 8.450 \\
\multirow[c]{9}{*}{3} & \multirow[c]{9}{*}{vm java} & java aot & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 43.5%, transparent 43.5%) 4.595 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 41.8%, transparent 41.8%) 4.638 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 40.7%, transparent 40.7%) 4.978 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 40.0%, transparent 40.0%) 4.825 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 38.7%, transparent 38.7%) 4.720 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 36.3%, transparent 36.3%) 4.431 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 33.9%, transparent 33.9%) 4.228 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 34.1%, transparent 34.1%) 4.175 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 28.3%, transparent 28.3%) 3.564 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 15.3%, transparent 15.3%) 2.661 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.9%, transparent 0.9%) 3.458 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.7%, transparent 0.7%) 10.831 \\
 &  & java grpc g1gc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 5.1%, transparent 5.1%) 0.763 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.1%, transparent 4.1%) 0.767 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.765 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.769 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.771 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.774 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.768 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.763 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.766 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 0.783 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 1.760 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 5.420 \\
 &  & java grpc pgc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.8%, transparent 4.8%) 0.742 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.0%, transparent 4.0%) 0.752 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.755 & \colorgreen \width10em \height80% 0.737 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 0.751 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 0.750 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 0.750 & \colorgreen \width10em \height80% 0.740 & \colorgreen \width10em \height80% 0.741 & \colorgreen \width10em \height80% 0.767 & \colorgreen \width10em \height80% 1.699 & \colorgreen \width10em \height80% 5.268 \\
 &  & java grpc sgc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.7%, transparent 4.7%) 0.726 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 3.9%, transparent 3.9%) 0.748 & \colorgreen \width10em \height80% 0.737 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 0.747 & \colorgreen \width10em \height80% 0.741 & \colorgreen \width10em \height80% 0.742 & \colorgreen \width10em \height80% 0.747 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 0.754 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.774 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.4%, transparent 0.4%) 0.815 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 1.782 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 5.363 \\
 &  & java grpc she & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 5.3%, transparent 5.3%) 0.788 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.4%, transparent 4.4%) 0.797 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.5%, transparent 0.5%) 0.791 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.5%, transparent 0.5%) 0.785 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.5%, transparent 0.5%) 0.796 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.5%, transparent 0.5%) 0.791 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 0.782 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.4%, transparent 0.4%) 0.776 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.4%, transparent 0.4%) 0.783 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.793 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 1.779 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.1%, transparent 0.1%) 5.944 \\
 &  & java grpc zgc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 5.3%, transparent 5.3%) 0.791 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.5%, transparent 4.5%) 0.804 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.6%, transparent 0.6%) 0.797 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.6%, transparent 0.6%) 0.795 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.5%, transparent 0.5%) 0.793 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.5%, transparent 0.5%) 0.793 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.4%, transparent 0.4%) 0.787 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.4%, transparent 0.4%) 0.776 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.6%, transparent 0.6%) 0.796 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 0.796 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 1.733 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 5.562 \\
 &  & java micronaut & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 5.6%, transparent 5.6%) 0.815 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.8%, transparent 4.8%) 0.833 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.9%, transparent 0.9%) 0.835 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.9%, transparent 0.9%) 0.824 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.1%, transparent 1.1%) 0.857 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.8%, transparent 0.8%) 0.827 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.8%, transparent 0.8%) 0.834 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.8%, transparent 0.8%) 0.821 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.9%, transparent 0.9%) 0.835 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.8%, transparent 0.8%) 0.863 & nan & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.2%, transparent 0.2%) 7.134 \\
 &  & kotlin grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 9.5%, transparent 9.5%) 1.205 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 8.2%, transparent 8.2%) 1.184 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.1%, transparent 4.1%) 1.164 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.0%, transparent 4.0%) 1.151 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 3.9%, transparent 3.9%) 1.143 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 3.6%, transparent 3.6%) 1.107 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 3.7%, transparent 3.7%) 1.128 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 3.7%, transparent 3.7%) 1.115 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 3.8%, transparent 3.8%) 1.122 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 3.1%, transparent 3.1%) 1.155 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.5%, transparent 0.5%) 2.669 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.3%, transparent 0.3%) 7.594 \\
 &  & scala akka & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 24.2%, transparent 24.2%) 2.670 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 21.2%, transparent 21.2%) 2.523 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 24.5%, transparent 24.5%) 3.294 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 27.8%, transparent 27.8%) 3.582 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 28.3%, transparent 28.3%) 3.654 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 27.8%, transparent 27.8%) 3.568 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 25.5%, transparent 25.5%) 3.360 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 28.0%, transparent 28.0%) 3.562 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 27.3%, transparent 27.3%) 3.465 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 19.3%, transparent 19.3%) 3.159 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.1%, transparent 2.1%) 5.976 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.0%, transparent 2.0%) 20.122 \\
4 & vm beam & elixir grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 15.8%, transparent 15.8%) 1.837 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 15.0%, transparent 15.0%) 1.889 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 11.0%, transparent 11.0%) 1.879 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 10.8%, transparent 10.8%) 1.836 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 10.6%, transparent 10.6%) 1.833 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 10.4%, transparent 10.4%) 1.800 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 9.8%, transparent 9.8%) 1.756 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 10.0%, transparent 10.0%) 1.751 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 9.6%, transparent 9.6%) 1.704 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 7.2%, transparent 7.2%) 1.662 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.3%, transparent 1.3%) 4.385 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.8%, transparent 0.8%) 11.448 \\
\multirow[c]{6}{*}{5} & \multirow[c]{6}{*}{interpreted} & dart grpc & \colorgreen \width10em \height80% 0.259 & \colorgreen \width10em \height80% 0.343 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.0%, transparent 0.0%) 0.739 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 8.2%, transparent 8.2%) 1.576 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 49.7%, transparent 49.7%) 5.846 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 59.6%, transparent 59.6%) 6.796 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 59.3%, transparent 59.3%) 6.837 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 60.5%, transparent 60.5%) 6.837 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 59.1%, transparent 59.1%) 6.645 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 46.4%, transparent 46.4%) 6.520 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.5%, transparent 2.5%) 6.649 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.7%, transparent 0.7%) 10.768 \\
 &  & node grpc st & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 51.5%, transparent 51.5%) 5.394 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 49.5%, transparent 49.5%) 5.427 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 36.2%, transparent 36.2%) 4.517 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 46.8%, transparent 46.8%) 5.523 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 46.2%, transparent 46.2%) 5.491 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 47.5%, transparent 47.5%) 5.564 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 44.7%, transparent 44.7%) 5.330 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 44.3%, transparent 44.3%) 5.200 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 43.3%, transparent 43.3%) 5.067 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 34.1%, transparent 34.1%) 5.000 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.9%, transparent 1.9%) 5.567 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.5%, transparent 0.5%) 8.719 \\
 &  & node grpcjs st & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 52.9%, transparent 52.9%) 5.533 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 61.0%, transparent 61.0%) 6.611 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 58.5%, transparent 58.5%) 6.842 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 56.9%, transparent 56.9%) 6.553 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 73.4%, transparent 73.4%) 8.279 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 75.2%, transparent 75.2%) 8.379 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 71.7%, transparent 71.7%) 8.104 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 72.6%, transparent 72.6%) 8.053 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 72.6%, transparent 72.6%) 7.992 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 54.9%, transparent 54.9%) 7.578 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 2.6%, transparent 2.6%) 6.900 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 0.7%, transparent 0.7%) 10.644 \\
 &  & php grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 31.8%, transparent 31.8%) 3.431 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 30.7%, transparent 30.7%) 3.503 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.2%, transparent 26.2%) 3.467 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.4%, transparent 26.4%) 3.437 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.2%, transparent 26.2%) 3.436 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.3%, transparent 26.3%) 3.413 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.1%, transparent 26.1%) 3.425 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.5%, transparent 26.5%) 3.414 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 26.7%, transparent 26.7%) 3.412 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 21.4%, transparent 21.4%) 3.422 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.5%, transparent 1.5%) 4.721 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.8%, transparent 1.8%) 18.743 \\
 &  & python grpc & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 81.9%, transparent 81.9%) 8.416 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 76.7%, transparent 76.7%) 8.226 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 11.166 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 10.959 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 11.016 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 10.894 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 11.010 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 10.812 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 98.8%, transparent 98.8%) 10.608 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 76.3%, transparent 76.3%) 10.226 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.3%, transparent 4.3%) 10.263 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.3%, transparent 1.3%) 14.892 \\
 &  & ruby grpc & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 10.224 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 10.623 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 73.9%, transparent 73.9%) 8.441 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 97.5%, transparent 97.5%) 10.701 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 95.9%, transparent 95.9%) 10.599 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 97.0%, transparent 97.0%) 10.592 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 96.1%, transparent 96.1%) 10.612 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 99.0%, transparent 99.0%) 10.711 & \colorred \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 100.0%, transparent 100.0%) 10.731 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 80.8%, transparent 80.8%) 10.782 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 4.0%, transparent 4.0%) 9.623 & \width10em \height80% \backgroundlinear-gradient(90deg, #aed6f1 1.2%, transparent 1.2%) 14.042 \\
\end{tabular}